---
title: "Data Analysis"
author: "Ethan Cohen"
date: "20/01/2020"
output: html_document
---

```{r, echo=FALSE, include=F}
library(ggplot2)
library(dplyr)

```


```{r, echo=FALSE}
Species <- c("Felis catus", "Puma concol", "Otocolobus manul","Ancinonyx jubatus",
             "Lynx canadensis","Leopardus paradalis", "Caracal caracal", 
             "Pardofelis badia", "Panthera leo", "Neofelis nebulosa")

MData <- read.csv("measurements.csv")
MData <- MData %>% rename(IDs = X, Spe = Species, LiW = Limb.Width, UnW = unitsW, LiL = Limb.Length, UnL = unitsL)

for (i in 1:length(MData$UnW)){
  if (MData$UnW[i] == "mm"){
    MData$LiW[i] = MData$LiW[i]/10
    MData$UnW[i] = "cm"
  }
}

for (i in 1:length(MData$UnL)){
  if (MData$UnL[i] == "mm"){
    MData$LiL[i] = MData$LiL[i]/10
    MData$UnL[i] = "cm"
  }
}
MData <- mutate(MData, Volume = pi*(MData$LiW/2)^2*LiL)

qplot (x=LiW, y=LiL, data=MData, xlab="Limb Width (cm)", ylab="Limb Length (cm)", main="Limb size comparison among feline species", colour=I(rgb(0,0,1)),size = I(4), alpha =I(0.8), shape=I(16))


```

Plot 1. Depicts the limb length(cm) plotted against the limb width (cm) of 10 species within the taxonimic family of felinae.


```{r, echo=FALSE}
mean(MData$Volume)
Big_MData <- subset(MData, Volume >11620.54)
qplot(x=LiW, y=LiL, data=Big_MData,xlab="Limb Width (cm)", ylab="Limb Length (cm)", main="Limb size comparison among feline species which have an above average volume", colour=I(rgb(0,0,1)),size = I(4), alpha =I(0.8), shape=I(16))
```

Plot 1. Depicts the limb length(cm) plotted against the limb width (cm) of 2 species with an above average (>11620.54cm^2) limb volume the taxonimic family of felineae.


## Feline Species With a Below Average Limb Volume
Table 1. Displays the species of the family felineae with a below average (<11620.54cm^2) limb volume. The average was obtained from the limb volumes of the 10 species used to in plot 1.

Species | Volume (cm^2)
----|----
Felis catus | 432.55
Puma concol | 10495.96
Otocolobus manul | 18.40
Lynx canadensis | 2.70
Leopardus paradalis | 68.40
Caracal caracal | 383.90
Panthera leo | 4367.5
Neofelis nebulosa | 0.14


```{r, echo=FALSE}
#New column called Fur_Pattern
Furi <- rbinom(10,1,0.5) + 1
type <- c("Spots", "Stripes")
Fur_Pattern <- type[Furi]

MData <- mutate(MData, Fur_Pattern = Fur_Pattern)

MD_Fur <- group_by(MData,Fur_Pattern)

Table1 <- summarise(MD_Fur, Avg_Length=mean(LiL), Sd_Length=sd(LiL), Avg_Width=mean(LiW), d_Width=sd(LiW))

View(Table1)

```







